---
title: "Sample Dashboard"
author: "Abigail Allison"
format: dashboard
css: formatting.css
theme: flatly
embed-resources: true
editor: source
engine: knitr
---

```{r setup}
#| echo: false
library(echarts4r)
source("scripts/00_clean.R")
source("scripts/01_wrangling.R")
source("scripts/02_viz_helpers.R")
```

# Overview

## Row

#### {.sidebar}

This project originated from a coding assessment request to analyze an outbreak line list and deliver a situational report for epidemiologists.

The underlying data is a synthetic COVID-19 [dataset](https://github.com/appliedepi/epiRhandbook_eng/tree/master/data/covid_example_data) provided by the global non-profit organization Applied Epi.

All graphs were created using the R package `echarts4r`.

::: {.callout-tip title="Usage Tips"}

Hover your cursor over graph contents to see a **tooltip** for additional information. 

Use the interactive **zoom slider** under the x-axis to select a specific date range.

To enter **full screen** with any graph, click the expand icon on the bottom right.
:::

### Column

#### Row

```{r new_cases}
#| content: valuebox
#| title: "New Cases (Latest Week)"
list(
  icon = "virus",
  value = tail(grouped_overview$`Number of Cases`, n=1)
) 
```

```{r total_cases}
#| content: valuebox
#| title: "Total Cases"
list(
  icon = "lungs",
  value = format(nrow(covid_clean), big.mark=",")
)
```

```{r total_deaths}
#| content: valuebox
#| title: "Total Deaths"
total_deaths <- sum(covid_clean$died_covid == "Yes", na.rm = TRUE)
list(
  icon = "people",
  value = format(total_deaths, big.mark=",")
)

```

#### Row
```{r epi_curve_overview}
grouped_overview |> 
  e_charts(week_start) |> 
  e_bar(`Number of Cases`, bind = label) |> 
  e_tooltip(
    formatter = tooltip_weekly_cases
  ) |> 
  e_axis_labels(y = "Cases (N)") |> 
  e_legend(show = FALSE) |> 
  e_title(text = "COVID-19 Cases by Week", left = "center", subtext = "2020-2021") |> 
  e_datazoom(x_index = 0, type = "slider") 

```

# Cases

## Row

### Column

```{r case_demo}
# demographics
cases_demo |>
  e_charts(case_race) |>
  e_pie(
    Count,
    radius = c("50%", "70%"),
    name = "Cases"
  ) |>
  e_title(text = "COVID-19 Cases by Race", left = "center", subtext = "2020–2021") |>
  e_legend(show = FALSE) |>
  e_tooltip(
    formatter = tooltip_race_pie
  ) |>
  e_color(rev(flatly_pie))
```

### Column

#### Row {height="60%"}

```{r admissions}
# most zip codes ARE in Fulton County
# Fulton County population 1,075,000 (2022)
admissions <- covid_clean |> 
  filter(hospitalized=="Yes") |> 
  mutate(county_pop = 1075000) |> 
  mutate(week_start = lubridate::floor_date(pos_sampledt, "weeks", week_start = 1)) |> 
  group_by(week_start) |> 
  filter(!is.na(week_start)) |> 
  summarize(Admissions = n()) |> 
  mutate(population = 1075000) |> 
  mutate(admit_rate = (Admissions/population)*100000)

admissions |> 
  mutate(admit_rate = format(round(admit_rate, 2), nsmall=2)) |> 
  e_charts(week_start) |> 
  e_line(admit_rate, name = "Admission Rate per 100K Residents") |> 
  e_title(text = "COVID-19 Hospitalization Rate by Specimen Collection Week (2020-2021)", subtext = "Per 100K Fulton County Residents (2022 Census)", left="center") |> 
  e_legend(show=FALSE) |> 
  e_datazoom(type="slider") |> 
  e_tooltip()

```

#### Row {height="40%"}

```{r table}
# table: top 10 postal codes
fulton_postal_codes <- c(30004,30005,30009,30022,30023,30075,30076,30077,30097,30213,30268,30272,30291,30301,30302,30303,30304,30305,30306,30307,30308,30309,30310,30311,30312,30313,30314,30315,30316,30318,30320,30321,30324,30325,30326,30327,30328,30328,30330,30330,30331,30334,30336,30337,30337,30339,30342,30343,30344,30344,30347,30348,30349,30349,30353,30354,30354,30355,30357,30358,30358,30361,30364,30364,30369,30370,30371,30374,30376,30377,30378,30379,30392,30394,31106,31107,31126,31131,31139,31150,31156)

covid_clean |> 
  filter(case_zip %in% fulton_postal_codes) |> # would normally geocode
  group_by(case_zip) |> 
  summarize(Cases = n()) |> 
  arrange(desc(Cases)) |> 
  head(n=10) |> 
  DT::datatable(rownames = FALSE, colnames=c("Postal Code", "Cases(N)"), caption = htmltools::tags$caption(style = 'caption-side: top; text-align: center; color:black; font-size:115%', 'Postal Codes with Most Cases (2020-2021)'
  ))
```

# Deaths

```{r deaths_by_week}

# combine
grouped_overview <- grouped_overview |>
  left_join(covid_deaths, by = "week_start") |>
  mutate(
    # replace NA with 0
    `Number of Deaths` = tidyr::replace_na(`Number of Deaths`, 0L),
    death_rate = `Number of Deaths` / `Number of Cases`,
    cdc_week = epiweek(week_start),
    label = paste0("CDC Week ", cdc_week)
  )
  
grouped_overview |>
  e_charts(week_start) |>
  e_bar(`Number of Deaths`, bind = label) |>
  e_line(death_rate, y_index = 1, bind = label, name = "Case Fatality Rate") |>
e_tooltip(
  trigger = "axis",
  formatter = tooltip_deaths
) |>
  e_y_axis(index = 1, name = "Case Fatality Rate") |>
  e_y_axis(name = "Deaths (N)") |>
  e_legend(show = TRUE, top = "8%") |>
  e_title(text = "COVID-19 Deaths and Case Fatality Rate by Week", left = "center", subtext = "2020-2021") |>
  e_datazoom(x_index = 0, type = "slider")

```


