---
title: "COVID-19 Situational Dashboard"
author: "Abigail Allison"
format:
  dashboard:
    theme: flatly
    orientation: rows
    scrolling: false
css: formatting.css
embed-resources: true
editor: source
engine: knitr
---

```{r setup}
#| echo: false
#| message: false
#| warning: false
library(echarts4r)
source("scripts/00_clean.R")
source("scripts/01_wrangling.R")
source("scripts/02_viz_helpers.R")
```

# Overview

## Row

#### {.sidebar}

##### Reporting Status

**Latest complete week**<br>
`r latest_complete_week_label`

The week beginning `r format(as.Date(final_week_start), "%b %d, %Y")` is incomplete and excluded from headline metrics.

##### Data Timing

Cases use specimen collection date, admissions use admission date, and deaths use date of death.

Median reporting lag was `r round(reporting_lag_median)` days; recent counts may increase as delayed reports are received.

##### Scope

Synthetic confirmed COVID-19 line-list data from [Applied Epi](https://github.com/appliedepi/epirhandbook/tree/master/data/covid_example_data).

### Column

#### Row {height="22%"}

::: {.valuebox title="Cases" icon="virus" color="primary"}
`r format(latest_complete_metrics$cases, big.mark = ",")`

[`r latest_complete_metrics$cases_wow_label`]{.kpi-delta}
:::

::: {.valuebox title="Admissions" icon="hospital" color="warning"}
`r format(latest_complete_metrics$admissions, big.mark = ",")`

[`r latest_complete_metrics$admissions_wow_label`]{.kpi-delta}
:::

::: {.valuebox title="COVID-19 Deaths" icon="heart-pulse" color="danger"}
`r format(latest_complete_metrics$deaths, big.mark = ",")`

[`r latest_complete_metrics$deaths_wow_label`]{.kpi-delta}
:::

#### Row {height="78%"}

```{r weekly-event-trends}
weekly_events |>
  e_charts(week_label) |>
  e_bar(
    cases_complete,
    name = "Cases",
    y_index = 0,
    itemStyle = list(color = dashboard_colors$cases),
    barMaxWidth = 18
  ) |>
  e_bar(
    cases_incomplete,
    name = "Cases (incomplete week)",
    y_index = 0,
    itemStyle = list(color = dashboard_colors$muted),
    barMaxWidth = 18,
    barGap = "-100%"
  ) |>
  e_line(
    admissions,
    name = "Admissions",
    y_index = 1,
    symbol = "none",
    lineStyle = list(color = dashboard_colors$admissions, width = 2.5),
    itemStyle = list(color = dashboard_colors$admissions)
  ) |>
  e_line(
    deaths,
    name = "Deaths",
    y_index = 1,
    symbol = "none",
    lineStyle = list(color = dashboard_colors$deaths, width = 2.5),
    itemStyle = list(color = dashboard_colors$deaths)
  ) |>
  e_title(
    text = "Weekly Event Trends",
    #subtext = "Correct event dates | gray = incomplete final week",
    subtextStyle = list(fontSize = 11),
    left = "center"
  ) |>
  e_legend(top = 48) |>
  e_tooltip(trigger = "axis", formatter = tooltip_weekly_events) |>
  e_x_axis(
    name = "Week starting Monday",
    nameLocation = "middle",
    nameGap = 32,
    axisLabel = list(formatter = axis_date_formatter, hideOverlap = TRUE),
    axisLine = list(lineStyle = list(color = dashboard_colors$grid))
  ) |>
  e_y_axis(
    index = 0,
    name = "Cases",
    nameGap = 12,
    nameTextStyle = list(
      color = dashboard_colors$muted,
      fontSize = 11,
      fontWeight = 600
    ),
    splitLine = list(lineStyle = list(color = dashboard_colors$grid))
  ) |>
  e_y_axis(
    index = 1,
    name = "Admissions | Deaths",
    nameGap = 12,
    nameTextStyle = list(
      color = dashboard_colors$muted,
      fontSize = 11,
      fontWeight = 600
    ),
    splitLine = list(show = FALSE)
  ) |>
  e_datazoom(x_index = 0, type = "slider", height = 20, bottom = 2) |>
  e_grid(top = 92, right = 64, bottom = 72, left = 64)
```

# Severity by Age

## Row

### Column {width="72%"}

```{r severity-age-chart}
age_severity |>
  e_charts(age_group) |>
  e_bar(
    hospitalization_rate,
    name = "Hospitalized among known",
    bind = hospitalization_detail,
    itemStyle = list(color = dashboard_colors$admissions),
    barMaxWidth = 42
  ) |>
  e_bar(
    death_rate,
    name = "COVID-19 deaths among known",
    bind = death_detail,
    itemStyle = list(color = dashboard_colors$deaths),
    barMaxWidth = 42
  ) |>
  e_title(
    text = "Severity by Age",
    subtext = "Rates among records with known outcome status; hover for counts and completeness.",
    subtextStyle = list(fontSize = 11),
    left = "center"
  ) |>
  e_legend(top = 48) |>
  e_tooltip(trigger = "axis", formatter = tooltip_age_severity) |>
  e_x_axis(
    name = "Age group",
    nameLocation = "middle",
    nameGap = 32,
    axisLine = list(lineStyle = list(color = dashboard_colors$grid))
  ) |>
  e_y_axis(
    axisLabel = list(formatter = percent_axis_formatter),
    splitLine = list(lineStyle = list(color = dashboard_colors$grid))
  ) |>
  e_grid(top = 92, right = 28, bottom = 60, left = 68)
```

### Column {width="28%"}

#### Key Signal

Adults **65+** account for these shares of reported outcomes:

- **`r sprintf("%.1f%%", age_65_summary$case_share)`** of reported cases
- **`r sprintf("%.1f%%", age_65_summary$hospitalization_share)`** of reported hospitalizations
- **`r sprintf("%.1f%%", age_65_summary$death_share)`** of reported COVID-19 deaths

These are shares of reported events, not population risks.

#### Outcome Completeness

Outcome status is available for **`r sprintf("%.1f%%", overall_hospitalization_completeness)`** of records for hospitalization and **`r sprintf("%.1f%%", overall_death_completeness)`** for COVID-19 death. Rates are calculated among records with known status and should be interpreted cautiously.
