---
title: "Sample Dashboard"
author: "Abigail Allison"
format: dashboard
css: formatting.css
theme: flatly
embed-resources: true
editor: source
engine: knitr
---

```{r setup}
#| echo: false
pacman::p_load(tidyverse, readxl, janitor, here, echarts4r, stringr, dt)

# made project for this
covid_raw <- read_excel("data/covid_example_data.xlsx")

# tidy column names with janitor
covid_clean <- clean_names(covid_raw)

# deduplicate any identical rows
covid_clean <- covid_clean[!duplicated(covid_clean),] # same # rows, no duplicates

# drop "_false" endings from column names
colnames_clean <- gsub("_false", "", colnames(covid_clean))
colnames(covid_clean) <- colnames_clean
```

```{r visuals}
# 1. Epi Curve Overview
# n=122 missing positive sample dates
pos_sampledt_no_na <- covid_clean |> 
  select(pos_sampledt, case_gender) |> 
  filter(!is.na(pos_sampledt))

# grab first day of week from pos_sampledt
pos_sampledt_no_na$week_start <- lubridate::floor_date(pos_sampledt_no_na$pos_sampledt, "weeks", week_start = 1)

grouped_overview <- pos_sampledt_no_na |> 
  group_by(week_start) |> 
  summarize(`Number of Cases` = n())

grouped_overview$cdc_week <- epiweek(grouped_overview$week_start)
grouped_overview$label = paste0("CDC Week ", grouped_overview$cdc_week)

# 2. Cases
cases_demo <- covid_clean |> 
  group_by(case_race) |> 
  summarize(Count=n()) |> 
  mutate(proportion = (Count/sum(Count)*100)) |> 
  mutate(case_race = str_to_title(ifelse(is.na(case_race), "MISSING", case_race))) |> 
  arrange(proportion) 

# 3. Deaths
covid_deaths <- covid_clean %>% 
  select(died_covid, died_dt, starts_with("case_")) %>% 
  filter(died_covid == "Yes")

```

# Overview

## Row

#### {.sidebar}

This project originated from a coding assessment request to analyze an outbreak line list and deliver a situational report for epidemiologists.

The underlying data is a synthetic COVID-19 dataset provided by the global non-profit organization [Applied Epi](https://www.appliedepi.org).

All graphs were created using the R package `echarts4r`, which is user friendly for basics but requires JavaScript for customization.

Throughout this dashboard, "CDC Week" is used to refer to MMWR week.


::: {.callout-tip title="Usage Tips"}

Hover your cursor over graph contents to see a **tooltip** for additional information. 

Use the interactive **zoom slider** under the x-axis to select a specific date range.

To enter **full screen** with any graph, click the expand icon on the bottom right.
:::

### Column

#### Row

```{r new_cases}
#| content: valuebox
#| title: "New Cases (Latest Week)"
list(
  icon = "virus",
  #color = "info",
  value = tail(grouped_overview$`Number of Cases`, n=1)
) 
```

```{r total_cases}
#| content: valuebox
#| title: "Total Cases"
list(
  icon = "lungs",
  #color = "danger",
  value = format(nrow(covid_clean), big.mark=",")
) # add new cases, deaths, R0?
```

```{r total_deaths}
#| content: valuebox
#| title: "Total Deaths"
list(
  icon = "people",
  #color = "danger",
  value = format(nrow(covid_deaths), big.mark=",")
) 
```

#### Row
```{r epi_curve_overview}
grouped_overview |> 
  e_charts(week_start) |> 
  e_bar(`Number of Cases`, bind = label) |> 
  e_tooltip(
    formatter = htmlwidgets::JS("
      function(params){
        return('<strong>' + params.name + 
                '</strong><br />Week Starting: ' + params.value[0] + 
                '<br />Number of Cases: ' + params.value[1]) 
                }
    ")
  ) |> 
  e_axis_labels(y = "Cases (N)") |> 
  e_legend(show = FALSE) |> 
  e_title(text = "COVID-19 Cases by CDC Week", left = "center", subtext = "2020-2021") |> 
  e_datazoom(x_index = 0, type = "slider") 

```

# Cases

## Row

### Column

```{r case_demo}
# load package for color palettes
# library(RColorBrewer)
# set vector to # of races present in data (n=8)
# my_colors <- rev(brewer.pal(n = 8, name = "Set1"))
# Added 'rev' to reverse color vector for Set1, otherwise not aesethetic on plot
flatly_pie <- c(
  "#4E79A7",  # muted blue (anchor)
  "#76B7B2",  # teal
  "#59A14F",  # green
  "#EDC948",  # soft gold (NOT bright yellow)
  "#AF7AA1",  # muted purple
  "#F28E2B",  # softened orange
  "#9C755F",  # warm gray-brown
  "#BAB0AC"   # neutral gray
)


# demographics
cases_demo |> 
  mutate(proportion = format(round(proportion, 2), nsmall=2)) |> 
  #arrange(desc(proportion)) |>
  e_charts(case_race) |> 
  e_pie(proportion, radius = c("50%", "70%"), name= "Percentage(%)") |>
  e_title(text = "COVID-19 Cases by Race", left="center", subtext = "2020-2021") |> 
  e_legend(show=FALSE) |> 
  e_tooltip() |> 
  e_color(rev(flatly_pie))
```

### Column

#### Row {height="60%"}

```{r admissions}
# most zip codes ARE in Fulton County
# Fulton County population 1,075,000 (2022)
admissions <- covid_clean |> 
  filter(hospitalized=="Yes") |> 
  mutate(county_pop = 1075000) |> 
  mutate(week_start = lubridate::floor_date(pos_sampledt, "weeks", week_start = 1)) |> 
  group_by(week_start) |> 
  filter(!is.na(week_start)) |> 
  summarize(Admissions = n()) |> 
  mutate(population = 1075000) |> 
  mutate(admit_rate = (Admissions/population)*100000)

admissions |> 
  mutate(admit_rate = format(round(admit_rate, 2), nsmall=2)) |> 
  e_charts(week_start) |> 
  e_line(admit_rate, name = "Admission Rate") |> 
  e_title(text = "COVID-19 Hospitalizations by Specimen Collection Week (2020-2021)", subtext = "Per 100K Fulton County Residents (2022 Census)", left="center") |> 
  e_legend(show=FALSE) |> 
  e_datazoom(type="slider") |> 
  e_tooltip()
```

#### Row {height="40%"}

```{r table}
# table: top 10 postal codes
fulton_postal_codes <- c(30004,30005,30009,30022,30023,30075,30076,30077,30097,30213,30268,30272,30291,30301,30302,30303,30304,30305,30306,30307,30308,30309,30310,30311,30312,30313,30314,30315,30316,30318,30320,30321,30324,30325,30326,30327,30328,30328,30330,30330,30331,30334,30336,30337,30337,30339,30342,30343,30344,30344,30347,30348,30349,30349,30353,30354,30354,30355,30357,30358,30358,30361,30364,30364,30369,30370,30371,30374,30376,30377,30378,30379,30392,30394,31106,31107,31126,31131,31139,31150,31156)

covid_clean |> 
  filter(case_zip %in% fulton_postal_codes) |> # would normally geocode
  group_by(case_zip) |> 
  summarize(Cases = n()) |> 
  arrange(desc(Cases)) |> 
  head(n=10) |> 
  DT::datatable(rownames = FALSE, colnames=c("Postal Code", "Cases(N)"), caption = htmltools::tags$caption(style = 'caption-side: top; text-align: center; color:black; font-size:115%', 'Postal Codes with Most Cases (2020-2021)'
  ))
  #cols_label(case_zip = "Postal Code", Cases = "Total (N)") |> 
  #tab_header(title = "Postal Codes with Most Cases (2020-2021)") |> 
  #cols_align(align = "center")
```

# Deaths

```{r deaths_by_week}

# add line for deaths
deaths <- covid_clean |> 
  filter(!is.na(pos_sampledt)) |> 
  filter(died == "Yes") |> 
  select(pos_sampledt)
  
deaths$week_start = lubridate::floor_date(deaths$pos_sampledt, "weeks", week_start = 1)
  
deaths <- deaths |> 
  group_by(week_start) |> 
  summarize(`Number of Deaths` = n()) 

deaths$cdc_week = epiweek(deaths$week_start)
deaths$label = paste0("CDC Week ", deaths$cdc_week)

# join deaths and grouped_overview
grouped_overview <- left_join(grouped_overview, deaths) # will join by week_start, common variable
# change NAs in number of deaths column to 0
grouped_overview[is.na(grouped_overview$`Number of Deaths`),5] <- 0
grouped_overview$death_rate = (grouped_overview$`Number of Deaths`/grouped_overview$`Number of Cases`)

deaths <- left_join(deaths, grouped_overview)

deaths |>
  e_charts(week_start) |>
  e_bar(`Number of Deaths`, bind = label) |>
  e_line(death_rate, y_index = 1, bind = label, name = "Case Fatality Rate") |>
e_tooltip(
  trigger = "axis",
  formatter = htmlwidgets::JS("
    function(params){

      function getY(p){
        var v = p.value;
        if (Array.isArray(v)) return v[v.length - 1]; // y
        return v;
      }

      var bar  = params.find(p => p.seriesType === 'bar');
      var line = params.find(p => p.seriesType === 'line');

      // CDC Week label comes from bind -> params[i].name
      var cdcWeekLabel = params[0].name; // e.g., 'CDC Week 31'

      // Week start date from axis value or from value[0]
      var weekStart =
        (params[0].axisValueLabel || params[0].axisValue) ||
        (Array.isArray(params[0].value) ? params[0].value[0] : params[0].name);

      var deaths = bar  ? getY(bar)  : null;
      var cfr    = line ? getY(line) : null;

      var deathsTxt = (deaths === null || deaths === undefined) ? 'NA' : deaths;

      var cfrTxt = (cfr === null || cfr === undefined || isNaN(cfr))
        ? 'NA'
        : (Number(cfr) * 100).toFixed(2) + '%';

      return '<strong>' + cdcWeekLabel + '</strong>'
        + '<br/>Week Starting: ' + weekStart
        + '<br/>Number of Deaths: ' + deathsTxt
        + '<br/>Case Fatality Rate: ' + cfrTxt;
    }
  ")
) |>
  e_y_axis(index = 1, name = "Case Fatality Rate") |>
  e_y_axis(name = "Deaths (N)") |>
  e_legend(show = TRUE, top = "8%") |>
  e_title(text = "COVID-19 Deaths and Case Fatality Rate by CDC Week", left = "center", subtext = "2020-2021") |>
  e_datazoom(x_index = 0, type = "slider")

```


